\sys: Automated Data-Parallel Shell Processing

### 2020-05-03 pldi/draft.tex
Refactor implementation nuggets and leave comment

@@

\TODO{Problem: Cat (and other commands that have multiple inputs) is
  not eager. Illustration: Show a parallel pipeline having a cat that
  gathers input somewhere (and a sort merge maybe). The way pipes
  work, this graph will essentially evaluate almost completely
  sequentially. Maybe go through the steps that cat asks for the first
  line of its first input until ithe first input is done and then goes
  to the second. This leads to processor under-utilization. Solution:
  We add eager nodes that eagerly consume all their input (even if
  their output does not ask for it). Eager nodes are themselves part
  of the dataflow graph (and model) and essentially perform an
  identity function. This makes all processors as utilized as possible
  and for CPU heavy tasks it leads to speedup. For IO heavy (and CPU
  light) tasks it has some overhead. We evaluate eager in section X
  (We should also say something about a CPU light task). A good thing
  is that they work for all possible commands with almost no overhead,
  without having to make each specific command eager.}

\TODO{(This problem is a nice followup for the one above) Problem:
  Dataflow graph termination. Illustration: Show the example of head
  deadlock and explain that this is inherent behaviour in the shell.
  (Solution: emptying pipelines or terminating processes) In order for
  things to terminate gracefully, dish wraps all commands in order to
  always consume all of their input by sending it down to
  /dev/null. Warning: This means that all commands in the pipeline
  must terminate gracefully if all of their output is consumed (if a
  process dies unexpectedly, noone will consume its input. We leave
  addressing that and other monitoring questions for future work). We
  evaluate the overhead of emptying all outputs in section X.}

\TODO{Problem: Pipeline width shrinks after parallel pure
  commands. Solution: Split node that resplits its input to multiple
  files in sequence (to preserve the model semantics). We should
  probably evaluate that too.}

\paragraph{Merge and Split Implementation}
%
It is common in the dataflow graph for a node to have many incoming
edges or many outgoing edges (when all but the last outgoing edges
are bounded). However some nodes, might only read their input from one
source (e.g. stdin). In this case, the distributed implementation
needs to concatenate input files (or conversely split output
files).
%
%% \tr{Show an example graph! of cat | tr | sort. The command in the
%%   middle must only take input from one file (maybe just stdin).}
%
We address this issue using primitive shell constructs for file
manipulation. More concretely a merger can be implemented by means of a simple
\ttt{cat}. Given a set of input files \ttt{in1, in2, ...} and an
output file \ttt{out}, a merger can be implemented as:

\begin{lstlisting}[language=sh, numbers=none]
 cat $in1 $in2 ... > $out
\end{lstlisting}

\noindent
On the other hand, given an input file \ttt{in} that is a FIFO pipe,
and two output files \ttt{out1, out2} the first of which is bounded
to \ttt{|N} lines, a 2-splitter can be implemented as:

\begin{lstlisting}[language=sh, numbers=none]
  tee >(head -n $N > $temp1;
        dd of=/dev/null > /dev/null 2>&1 &
        cat $temp1 > $out1) |
       (tail -n +${N+1} > $out2;
        dd of=/dev/null > /dev/null 2>&1)
\end{lstlisting}
 %% head -n $N $in > $out1 ; cat $in > $out2

\noindent
Using the 2-splitter as a building block, a pipe can be split to
arbitrarily many different pipes.

\heading{Dangling FIFOs and Zombie Producers}
% \nv{I think all of these titles should be the problem/challenge, not the solution}
% A process writing to a pipe needs to be notified when its consumer exits early before consuming all its output.
Under normal operation, a producer commonly exits after it has produced and sent all of its output to a pipe.
% \ttt{SIGPIPE} 
However, if its consumer exits early, the producer needs to be notified so that it stops appending to the stream. % ---so that it avoids unnecessary work and buffering challenges in the kernel.

@@

To solve this problem, \sys emits cleanup logic that operates from the end of the pipeline and towards its start.
The emitted code first gathers the ID of the last process and passes it as a parameter to \ttt{wait}; 
  this causes \ttt{wait} to block only on the last consumer of the pipeline.
Right after \ttt{wait}, \sys inserts a routine that delivers \ttt{PIPE} signals to any remaining processes upstream.
This epilogue is repeated for every \todo{merge/join} point in the transformed pipeline.
% Earlier: need to define split (fork?) and merge (join?) points; this should be "up to next merge point upstream"

>>>>

\heading{Overcoming Laziness}

\noindent
\TODO{Problem: Cat (and other commands that have multiple inputs) is
  not eager. Illustration: Show a parallel pipeline having a cat that
  gathers input somewhere (and a sort merge maybe). The way pipes
  work, this graph will essentially evaluate almost completely
  sequentially. Maybe go through the steps that cat asks for the first
  line of its first input until ithe first input is done and then goes
  to the second. This leads to processor under-utilization. Solution:
  We add eager nodes that eagerly consume all their input (even if
  their output does not ask for it). Eager nodes are themselves part
  of the dataflow graph (and model) and essentially perform an
  identity function. This makes all processors as utilized as possible
  and for CPU heavy tasks it leads to speedup. For IO heavy (and CPU
  light) tasks it has some overhead. We evaluate eager in section X
  (We should also say something about a CPU light task). A good thing
  is that they work for all possible commands with almost no overhead,
  without having to make each specific command eager.}

@@

\heading{Readjusting Graph Width}

\noindent
\TODO{Problem: Pipeline width shrinks after pure commands. Solution:
  Split node that resplits its input to multiple files in sequence (to
  preserve the model semantics).}

%% Old and obsolete

%% \noindent
%% It is common in the dataflow graph for a node to have many incoming
%% edges or many outgoing edges (when all but the last outgoing edges
%% are bounded). However some nodes, might only read their input from one
%% source (e.g. stdin). In this case, the distributed implementation
%% needs to concatenate input files (or conversely split output
%% files).
%
%% \tr{Show an example graph! of cat | tr | sort. The command in the
%%   middle must only take input from one file (maybe just stdin).}
%
%% We address this issue using primitive shell constructs for file
%% manipulation. More concretely a merger can be implemented by means of a simple
%% \ttt{cat}. Given a set of input files \ttt{in1, in2, ...} and an
%% output file \ttt{out}, a merger can be implemented as:

%% \begin{lstlisting}[language=sh, numbers=none]
%%  cat $in1 $in2 ... > $out
%% \end{lstlisting}

%% \noindent
%% On the other hand, given an input file \ttt{in} that is a FIFO pipe,
%% and two output files \ttt{out1, out2} the first of which is bounded
%% to \ttt{|N} lines, a 2-splitter can be implemented as:

%% \begin{lstlisting}[language=sh, numbers=none]
%%   tee >(head -n $N > $temp1;
%%         dd of=/dev/null > /dev/null 2>&1 &
%%         cat $temp1 > $out1) |
%%        (tail -n +${N+1} > $out2;
%%         dd of=/dev/null > /dev/null 2>&1)
%% \end{lstlisting}
 %% head -n $N $in > $out1 ; cat $in > $out2

%% \noindent
%% Using the 2-splitter as a building block, a pipe can be split to
%% arbitrarily many different pipes.

\TODO{Talk about eagerness in split and how we needed to add eager
  nodes to n-1 outputs.}

\heading{Dangling FIFOs and Zombie Producers}

%% \TODO{(This problem is a nice followup for the one above) Problem:
%%   Dataflow graph termination. Illustration: Show the example of head
%%   deadlock and explain that this is inherent behaviour in the shell.
%%   (Solution: emptying pipelines or terminating processes) In order for
%%   things to terminate gracefully, dish wraps all commands in order to
%%   always consume all of their input by sending it down to
%%   /dev/null. Warning: This means that all commands in the pipeline
%%   must terminate gracefully if all of their output is consumed (if a
%%   process dies unexpectedly, noone will consume its input. We leave
%%   addressing that and other monitoring questions for future work). We
%%   evaluate the overhead of emptying all outputs in section X.}

% \nv{I think all of these titles should be the problem/challenge, not the solution}
% A process writing to a pipe needs to be notified when its consumer exits early before consuming all its output.
Under normal operation, a producer commonly exits after it has produced and sent all of its output to a pipe.
% \ttt{SIGPIPE} 
However, if its consumer exits early, the producer needs to be notified so that it stops appending to the stream. % ---so that it avoids unnecessary work and buffering challenges in the kernel.

@@

To solve this problem, \sys emits cleanup logic that operates from the
end of the pipeline and towards its start.  The emitted code first
gathers the IDs of the output processes and passes them as parameters
to \ttt{wait}; this causes \ttt{wait} to block only on the output
producers of the dataflow graph.  Right after \ttt{wait}, \sys inserts
a routine that delivers \ttt{PIPE} signals to any remaining processes
upstream.  This epilogue is repeated for every \todo{merge/join} point
in the transformed pipeline \kk{I don't understand the final sentence}.
% Earlier: need to define split (fork?) and merge (join?) points; this should be "up to next merge point upstream"
